---
title: "Data Processing and Visulization of  of COVID-19 Outbreak in EU Countries"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
    css: "COVID-19-IT.css"
---

# Package Initialization

Here, we use the the *Coronavirus* package develpped and maintained with *Rami Krispin* at this link:

https://github.com/RamiKrispin/coronavirus

In addition, the *Tidyverse* package also was needed to manipulate and viulize the data.

```{r warning=FALSE, include=FALSE}
# Install the coronavirus package

if ("coronavirus" %in% rownames(installed.packages()))
{
  library(coronavirus)
  data("coronavirus")
} else {
  devtools::install_github("RamiKrispin/coronavirus")
  library(coronavirus)
  data("coronavirus")
}

# Install the Tidyverse package

if ("tidyverse" %in% rownames(installed.packages()))
{
  library(tidyverse)
} else {
  install.packages("tidyverse")
  library(tidyverse)
}
```


# Data Manupilation

Here, the dplyr was used to selcet the following countries:
+ Italy
+ Soth Korea
+ France
+ Denmark
+ Norway
+ Sweden

Then, we caculated the sum of the whole confirmed cases until the given time. In addition, the "row_number" was added to find the days after the Corona Cases has been reached above 25 cases.


```{r}
eu_cases <- coronavirus %>% 
  group_by(Country.Region, date) %>% 
  summarise(Cases=sum(cases)) %>% 
  filter(Country.Region== "Italy" | Country.Region== "Korea, South" | Country.Region== "France" | Country.Region== "Denmark" | 
           Country.Region== "Norway" |
           Country.Region=="Sweden") %>%
  group_by(Country.Region) %>% 
  mutate(tot_confirmed = cumsum(Cases)) %>% 
  filter(Cases > 25) %>% 
  mutate(id = row_number()) %>% 
  filter(id < 22)
```

# Mutating Countries Population to the Data

Here, the population of the countries were found and corrosponding population was added to the column "pop" in the dataset.

```{r}
populations <- c(81.16, 82.79, 60.48, 51.47, 66.44, 5.6, 5.36, 1386, 67, 10 )
names(populations) <- c("Iran", "Germany", "Italy",  "Korea, South" 
                        ,"UK", "Denmark", "Norway" ,"Mainland China", "France", "Sweden")

eu_cases$pop <- rep(0,length(eu_cases$Country.Region))

for (i in 1:length(eu_cases$Country.Region)) {
  country <- as.character(eu_cases$Country.Region[i])
  eu_cases$pop[i] <- populations[[country]]
}  
```

# ggplot Theme

Great thanks to Cedric Scherer (cedricphilippscherer@gmail.com) for his beatiful theme!, this theme was used as the theme of the ggplot in the bottom

```{r}
source(here::here("theme", "tidy_grey.R"))
```


If you are using the windows, you must install the following the fonts Poppins and Roboto Mono and then:


```{r}
windowsFonts("Roboto Mono" = windowsFont("Roboto Mono"))
windowsFonts("Poppins" = windowsFont("Poppins"))
```

# Plot Size Modification

Here, we use the costum edit in order to plot the ggplot output in the longer format than the defult. 

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center",
               fig.width = 10, fig.height = 14,
               dev.args = list(pointsize=10),
               out.width = "90%", dpi = 300,
               cache = FALSE,
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE, message = FALSE)

```


# Visulization:

```{r}
eu_cases <- eu_cases %>% mutate(tot_confirmed_per = tot_confirmed/pop)

eu_cases$Country.Region = factor(eu_cases$Country.Region, 
                              levels=c("Italy", "Korea, South","Norway","France" ,"Denmark" 
                                        ,"Sweden"))

ggplot(eu_cases, aes(id, tot_confirmed_per, fill = Country.Region)) +
    geom_area() +
    geom_area(alpha = 0.75) +
    facet_wrap(~ Country.Region, nrow = 4) + 
    theme(legend.position = "none",
          plot.title = element_text(size=12),
          axis.title = element_text(size = 12, face = "plain"),
          axis.title.x = element_text(size = 12, face = "plain"),
          axis.title.y = element_text(size = 12, face = "plain"),
          axis.text = element_text(family = "Roboto Mono", size = 8.5),
          axis.ticks.x = element_line(colour = "grey40", size = 1),
          axis.ticks.y = element_line(colour = "transparent", size = 0.3),
          panel.grid.major.y = element_line(colour = "grey40", size = 0.3),
          panel.border = element_blank(),
          strip.background = element_rect(color = "transparent"),
          strip.text = element_text(size = 13, colour = "white"),
          panel.spacing = unit(15, "pt")) +
  scale_x_continuous(limits = c(1, 21)) +
    labs(x = "Days after Totoal Confirmed Caes > 25", y = "Total Cases Per Million Inhabitants\n", 
         title = "Development of Total Confirmed Cases\nPer Million inhabitants",
         caption = "\nVisualization by Peyman Kor  |  Data source: JHU & WHO")

```

